---
title: "Gender Differences in Attitudes Toward Having Children Among Childless U.S. Adults"
author: "Grace Williams"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Introduction

I am interested in exploring data related to childcare.

My research question is: How does gender relate to U.S. adults’ intentions or feelings about having children among those who have never been parents? I want to find out if women and men differ in how they describe their thoughts about having children. My hypothesis is that women are more likely than men to say that they “never wanted to have children” or “weren’t sure” compared to men, reflecting social and behavioral patterns in intention of raising kids. The key independent variable is F_GENDER, which indicates whether the respondent self-reported as a woman, a man, in some other way, or refused to answer. The dependent variable is the survey question (KIDSEVER_W147): “Which of the following best describes your thoughts on having children?” with response options: "There was a time when you wanted to have children," "You never wanted to have children," "You weren’t sure whether or not you wanted to have children," and "Refused." My hypothesis would be supported if the responses differ significantly by gender, specifically if women were more likely to respond with "You never wanted to have children" and "You weren’t sure whether or not you wanted to have children." The hypothesis would be disproved if there is no difference by gender, or if men are more likely than women to report they never wanted children or were unsure.

## Data and Measures

This analysis uses data from the Pew Research Center American Trends Panel Wave 147, which surveys U.S. adults who have never had children. The data are cross-sectional, meaning they were collected at one point in time rather than over a longer period. The main independent variable is gender, measured using the variable `F_GENDER`, which I coded to include only male and female respondents. The dependent variable comes from the question “Which of the following best describes your thoughts on having children?” (`KIDSEVER_W147`). The responses include whether someone once wanted children, never wanted children, or was unsure. I cleaned the data by removing missing and refused responses and coding the variables into more readable categories. My hypothesis is supported if women are significantly more likely than men to report never wanting children or being unsure, so if there is a positive and statistically significant coefficient for male respondents in the regression model predicting the binary outcome of non-parental certainty. If the coefficient is not statistically significant or is close to zero, I would conclude that there is no meaningful gender difference.

```{r, echo=FALSE}
library(tidyverse)
library(knitr)
library(broom)

panel_data <- read_csv("ATP W147.csv")
head(panel_data)
clean_data <- panel_data |>
  select(F_GENDER, KIDSEVER_W147, F_AGECAT) |>
  rename(
    gender = F_GENDER,
    child_views = KIDSEVER_W147,
    age_group = F_AGECAT
  ) |>
  filter(gender %in% c(1, 2)) |>
  filter(!is.na(child_views)) |>
  filter(child_views != "Refused") |>
  mutate(
    gender = factor(gender, labels = c("Male", "Female")),
    child_views = factor(child_views,
      levels = c(1, 2, 3),
      labels = c(
        "Wanted children at one point",
        "Never wanted children",
        "Unsure")))
```
```{r, echo=FALSE}
ggplot(data = clean_data, mapping = aes(x = child_views, fill = gender)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Attitudes Toward Having Children Among Childless U.S. Adults",
    subtitle = "Among U.S. adults who have never had children",
    x = "Response",
    y = "Number of respondents",
    fill = "Gender",
    caption = "Source: Pew Research Center ATP Wave 147"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r, echo=FALSE}
summary_table <- clean_data |>
  group_by(gender, child_views) |>
  summarize(n = n(), .groups = "drop") |>
  mutate(percent = n / sum(n) * 100) 
kable(summary_table, digits = 2)
```

```{r, echo=FALSE}
ggplot(data = summary_table, mapping = aes(x = child_views, y = percent, fill = gender)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Percentage of Responses by Gender",
    x = "Views on Having Children",
    y = "Percent of respondents (%)",
    fill = "Gender"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## Empirical Results

This analysis compares how men and women differ in their responses to the question about having children. The results show that women are slightly more likely than men to say that they wanted children at one point, while men are somewhat more likely to report that they never wanted children or were unsure. However, the differences between the groups are relatively small. Overall, this suggests that gender is associated with attitudes toward having children, but the relationship is not particularly strong. These findings provide limited support for my hypothesis that gender influences how people think about having children. 

```{r, echo=FALSE}
clean_data <- clean_data |>
  mutate(
    no_kids_binary = ifelse(
      child_views %in% c("Never wanted children", "Unsure"),
      1,
      0
    )
  )
model_binary <- lm(no_kids_binary ~ gender, data = clean_data)
kable(tidy(model_binary), digits = 3)
```

To better analyze this relationship, I created a binary outcome variable that groups respondents who said they “never wanted children” or were “unsure” into one category, and those who said they “wanted children at one point” into another. The regression results show that the intercept is 0.576, meaning that about 57.6% of men in the sample report that they never wanted children or were unsure. The coefficient for genderFemale is -0.051, indicating that women are approximately 5.1 percentage points less likely than men to report never wanting children or being unsure. This difference is statistically significant (p = 0.011), meaning we have evidence to reject the null hypothesis of no difference between men and women at the 5% significance level. The standard error of 0.020 suggests the estimate is relatively precise. The standard errors are relatively small, and the p-value indicates that the gender coefficient is statistically significant at conventional levels (p < 0.05). However, because this is cross-sectional observational data, this relationship should not be interpreted causally. This could be due to confounding variables such as age, education, relationship status, and religion, which may influence both gender differences and attitudes toward having children.

```{r, echo=FALSE}
model_age <- lm(no_kids_binary ~ gender + age_group, data = clean_data)
kable(tidy(model_age), digits = 3)
```
To assess whether age differences explain the relationship between gender and attitudes toward having children, I used a regression model that includes both gender and age group, showing that the coefficient for gender remains negative and statistically significant (estimated coefficient = -0.052, p = 0.009). Therefore, women are still about 5.2 percentage points less likely than men to report never wanting children or being unsure, even after accounting for age. In contrast, age group is not statistically significant (estimated coefficient = 0.023, p = 0.243), indicating that differences in age groups do not strongly predict the outcome in this model. Overall, this indicates that the observed gender difference is not explained by age differences in the sample. However, because this is still observational cross-sectional data, the results cannot be interpreted causally, and other unobserved factors may still influence both gender and attitudes toward having children.

## Conclusion
Overall, the results show that there are some small differences between men and women in how they think about having children, but the relationship is not very strong. Women are slightly more likely to say they once wanted children, while men are a bit more likely to say they never wanted children or are unsure. This provides only limited support for my hypothesis. The results do not support my original hypothesis; instead, they suggest the opposite pattern. There are a few limitations to this analysis. Because the data are cross-sectional, it is not possible to make causal claims. There may also be other factors, like socioeconomic status or relationship status, that affect both gender and attitudes toward having children. Additionally, missing or refused responses could affect the results. If I had more time, I would try to include more variables in the analysis and explore how other factors might interact with gender.

```{r print-code, ref.label=knitr::all_labels(), echo = TRUE, eval=FALSE}

```